### 2018-08-06 supp_material.tex
Supp Info

\documentclass[3p,onecolumn]{elsarticle}

\usepackage{natbib}
\usepackage[version=4]{mhchem}
\usepackage{graphicx,bm}
\journal{Journal of \LaTeX\ Templates}

\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Supporting material for ``A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}''}

%% Group authors per affiliation:
\author[rvt]{A.J.~Silveira}
\author[rvt]{S.~Pereda}
\author[focal,els]{F.W.~Tavares}
\author[focal]{C.R.A.~Abreu\corref{cor1}}
\ead{abreu@eq.ufrj.br}

\address[rvt]{Planta Piloto de Ingenier\'ia Qu\'imica, PLAPIQUI, Universidad Nacional del Sur,Camino La Carrindanga Km 7-CC: 717, Bah\'ia Blanca, Argentina}
\address[focal]{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro,Rio de Janeiro, RJ 21941-909, Brazil}
\address[els]{COPPE, Universidade Federal do Rio de Janeiro, Rio de Janeiro, RJ 21941-909, Brazil}

\cortext[cor1]{Corresponding author}
\end{frontmatter}

\begin{figure}[ht]
\includegraphics[]{rdf-koller}
\caption{\ce{[emim][B(CN)_4]}}
\label{fig:pressure}
\end{figure}

\begin{figure}[ht]
\includegraphics[]{rdf-Batista.pdf}
\caption{\ce{[emim][B(CN)_4]}}
\label{fig:pressure}
\end{figure}

\begin{figure}[ht]
\includegraphics[]{rdf-Liu.pdf}
\caption{\ce{[emim][B(CN)_4]}}
\label{fig:pressure}
\end{figure}

\begin{figure}[ht]
\includegraphics[]{rdf-Weber.pdf}
\caption{\ce{[emim][B(CN)_4]}}
\label{fig:pressure}
\end{figure}

\end{document}